Fast Sorting on Modern Computers

### 2005-08-31 docs/tech_report/conclusions.tex
spell checked

\label{conclusions}
This section presents a summary of the results from previous chapters, lists
important branch prediction results and discoveries, compares our results with
LaMarca's and discusses important cache results, discusses the best sort that we
found in a number of catagories, and suggests contributions made to computer
science by this research.

@@

In several cases, we found that direct-mapped caches performed better than
fully-associative ones, resulting in fewer misses. This was due to they're
replacement policy, and that the sorts in question were heavily optimized for
the direct-mapped caches. None of the changes slowed the fully-associatvive
caches however.

@@

In addition to the double alignement, we also added two new improvements. The
mergesorts were tiled for the level 1 caches as well as the level 2, slightly
reducing the number of level 1 cache misses, and we removed the need for a step
to copy the array back from the auxilliary array; instead, we presort a step
further in cases where this occurs, and find the merge naturally ends in the
correct destination.

>>>>

\label{conclusions}
This section presents a summary of the results from previous chapters, lists
important branch prediction results and discoveries, compares our results with
LaMarca's and discusses important cache results, discusses the best sort that we
found in a number of categories, and suggests contributions made to computer
science by this research.

@@

In several cases, we found that direct-mapped caches performed better than
fully-associative ones, resulting in fewer misses. This was due to they're
replacement policy, and that the sorts in question were heavily optimized for
the direct-mapped caches. None of the changes slowed the fully-associative
caches however.

@@

In addition to the double alignment, we also added two new improvements. The
mergesorts were tiled for the level 1 caches as well as the level 2, slightly
reducing the number of level 1 cache misses, and we removed the need for a step
to copy the array back from the auxiliary array; instead, we presort a step
further in cases where this occurs, and find the merge naturally ends in the
correct destination.

### 2005-08-31 docs/tech_report/heap.tex
spell checked

The relationship between these numbers and $H_N-1$ is relatively straighforward.
$H_N-1$ represents the cumulative number of changes in the minimum. There first
element is guaranteed to be the minimum. The second element has a 50\% chance of
being smaller than the first. If we examine a third element, then the chances of
that element being smaller than two other elements is one in three. Therefore,
this branch will be taken one third of the time. As the number of elements we
examine increases, the probability of the nth element being smaller than all
elements seen so far is $\frac{1}{n}$.

@@

We investigated the trade-offs in these two effects and we discovered that using
a single branch does indeed cause the branches for the ealier comparisons to
become more predictable. However, this improvement is outweighed by a loss in
predictability of the later comparisons. The overall prediction accuracy fell by
around 2\% when we used a single branch, rather than unrolling the loop. This
result is interesting because it shows that relatively small changes in the way
that an algorithm is implemented can have an effect on branch prediction
accuracy. In this case, unrolling does not just remove loop overhead. It also
improves the predictability of the branches within the loop.

>>>>

The relationship between these numbers and $H_N-1$ is relatively straightforward.
$H_N-1$ represents the cumulative number of changes in the minimum. There first
element is guaranteed to be the minimum. The second element has a 50\% chance of
being smaller than the first. If we examine a third element, then the chances of
that element being smaller than two other elements is one in three. Therefore,
this branch will be taken one third of the time. As the number of elements we
examine increases, the probability of the nth element being smaller than all
elements seen so far is $\frac{1}{n}$.

@@

We investigated the trade-offs in these two effects and we discovered that using
a single branch does indeed cause the branches for the earlier comparisons to
become more predictable. However, this improvement is outweighed by a loss in
predictability of the later comparisons. The overall prediction accuracy fell by
around 2\% when we used a single branch, rather than unrolling the loop. This
result is interesting because it shows that relatively small changes in the way
that an algorithm is implemented can have an effect on branch prediction
accuracy. In this case, unrolling does not just remove loop overhead. It also
improves the predictability of the branches within the loop.
